---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## 1.0 Loading and preprocessing the data

```{r}

library(ggplot2)
library(dplyr)

# load data
data <- read.csv(unz("activity.zip", "activity.csv"))

# change date class as Date
data$date = as.Date(as.character(data$date))


```



## What is mean total number of steps taken per day?

```{r}


data_sum = data %>% group_by(date) %>% summarise(total_number_steps_taken_each_day = sum(steps)) 

data_sum

data_sum %>% ggplot(aes(x= total_number_steps_taken_each_day)) + geom_histogram(binwidth = 2000)


summary(data_sum$total_number_steps_taken_each_day)

```

The mean of the total number of steps taken per day is 10766 while the median is 10765



## What is the average daily activity pattern?

```{r}

# time series

data %>% mutate(steps = ifelse(is.na(steps),0,steps)) %>% group_by(interval) %>% summarise(average_number_steps_taken_each_day = mean(steps)) %>% ggplot(aes(x=interval,y=average_number_steps_taken_each_day)) + geom_line()


# 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps

data2 = data %>% mutate(steps = ifelse(is.na(steps),0,steps)) %>% group_by(interval) %>% summarise(mean_number_steps_taken_each_day = mean(steps))

data2 = data2 %>% arrange(desc(mean_number_steps_taken_each_day))

data2$interval[1]

```

The 835,  5-minute interval on average across all the data in dataset which contains the max number of steps


## Imputing missing values

```{r}


# Check missing data
summary(data)

# Impute missing data by exchanging NA into mean steps

data_no_na = data %>% filter(!is.na(steps))

meansteps = mean(data_no_na$steps)

data3 = data %>%  mutate(steps = ifelse(is.na(steps),meansteps,steps))

# no more na in data

summary(data3)


# plot histogram

data3 %>% group_by(date) %>% summarise(total_number_steps_taken_each_day = sum(steps)) %>% ggplot(aes(x= total_number_steps_taken_each_day)) + geom_histogram(binwidth = 2000)


# median and mean. no change

summary(data_sum$total_number_steps_taken_each_day)


```

The value does not differ before and after we impute the NA values into the data. However the total number of steps taken for each day increases.


## Are there differences in activity patterns between weekdays and weekends?

```{r}

data3$days = weekdays(data3$date)

data3 = data3 %>% mutate(days_group = ifelse(days == "Saturday" |days == "Sundays","weekends","weekdays"))

data3$days_group = as.factor(data3$days_group)

data3 %>% group_by(interval,days_group) %>% summarise(average_number_steps_taken_each_day = mean(steps)) %>% ggplot(aes(x=interval,y=average_number_steps_taken_each_day)) + geom_line() + facet_grid(~days_group)


```

Yes there are differences between weekends which has more average steps compare to weekdays which have slightly lower average steps